{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion}

### 2016-12-28 proposal/lipservice.tex
write about pintos

>>>>

The Pintos architecture \cite{pintos}, outlined in Section~\ref{sec:overview-pintos}, provides an opportunity to test Landslide's pedagogical mettle beyond CMU's walls.
For the experiments in Chapter~\ref{chap:quicksand}, I already implemented rudimentary support to test Pintos kernels.
Because Pintos provides substantially more basecode than Pebbles, most if not all of Landslide's kernel instrumentation can be done automatically,
using the names of the core scheduler functions already provided.

{\bf Stuff I already did.}
The existing Pintos support includes
a new set of annotations (e.g., {\tt mutex\_lock} in Pebbles is now called {\tt sema\_down};
handling several quirks of the basecode's scheduling behaviour (ask me if you really care to know);
extending the heap tracker to handle Pintos's page allocator {\tt palloc} as well as the kernel {\tt malloc},
as well as the fact that kernel memory is not direct-mapped like in Pebbles;
new thread-liveness code to detect when a test case terminates;
skipping some busy-wait loops in the boot sequence used for device communication,
a complicated fall-back case in Landslide's timer injection algorithm the workings of which I don't even remember anymore,
and extending the lock-set analysis to include Pintos's interrupt-disabling infrastructure (which was actually pretty tricky when using vector clocks for pure-HB).
All these features apply to behaviours included with the stock base-code, so should apply generally to all but the most adventurous student implementations.

{\bf Stuff I still need to do.}
However, as you might have guessed, there are still a few cases where the existing instrumentation is not fully general.
When preparing Chapter~\ref{chap:quicksand}'s experiments, I needed to adjust some student submissions to be compatible with Landslide's existing annotation process.
In order to distribute Landslide for student use beyond CMU, it will need to support these cases automatically.
These include:
\begin{itemize}
	\item Allowing for the priority scheduler's runqueues to be implemented as an array of queues, rather than the single queue head as provided in the basecode.
	\item Insert the {\tt tell\_landslide\_sleeping} annotation more intelligently than looking for a {\tt list\_insert\_ordered} call (a basecode function which most, but not all, students use).
	\item Some versions of the basecode distribute a {\tt sema\_up} implementation which {\tt yield}s unconditionally, which Landslide must bypass.
	\item Some students use {\tt timer\_sleep} or {\tt while (!flag) continue} when they should use {\tt yield}, which can lead to false-positive deadlocks if not automatically replaced.
	\item I'm sure I've forgotten some less-common ones, which I expect to handle on-the-fly when students email me Landslide crash reports.
\end{itemize}
When fixing my sample of Pintoses by hand, there were also myriad deterministic bugs, such as use-after-frees e.g.~arising from unsafe {\tt strlen} calls.
As with Pebbles student experiments, however, I intend the students to encounter such bug reports and fix them on their own, despite not being concurrency bugs.
It would be a Ph.D. unto itself to automatically ensure stable determinized execution of arbitrary student code.
Actually, probably outright impossible.

{\bf Experimental goals.}
Having only one semester (ideally; or two, if my timeline ends up delayed) to test Landslide with Pintos students,
I'll not be able to assemble as comprehensive a dataset as I have with 15-410 at CMU.
This project will serve more as a proof-of-concept, demonstrating that model checking can realistically replace stress testing in operating systems courses worldwide.
Nevertheless, I intend also to analyze what little data I can gather, comparing the incidences and types of concurrency bugs found between Pebbles and Pintos populations.
This will perhaps shed light on the advantages and/or shortcomings of either project,
leading to recommendations for improving either project's educational value (independently of using MC).
